{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-11-21 sftrends_manu.Rnw
new trends map with bar plots is done

<<trndmap, fig.height = 6, fig.width = 9, cache = T, out.width = '0.75\\textwidth', fig.cap = 'Percent changes in nitrogen analytes for (a) annual and (b) seasonal (monthly) periods in the record.  Changes are based on the difference between the ending and starting estimates for the flow-normalized estimates within each period.  Points are colored for direction (red increasing, green decreasing) and sized for relative magnitude. Station names are shown in the top left panel. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'>>=
data(mods)

mobrks <- list(c(3, 4, 5), c(6, 7, 8), c(9, 10, 11), c(12, 1, 2))
yrbrks <- c(1975, 1995, 2000, 2008, 2014)
molabs <- c('Spring', 'Summer', 'Fall', 'Winter')
yrlabs <- c('1976-1995', '1996-2000', '2001-2008', '2009-2014')

trnd_map(mods = mods, molabs = NULL, yrlabs = yrlabs, yrbrks = yrbrks, sz_rng = c(2, 8), leg_lab = '(a) Annual % change')
trnd_map(mods = mods, yrlabs = NULL, molabs = molabs, mobrks = mobrks, sz_rng = c(2, 8), stat_labs = F, leg_lab = '(b) Seasonal % change')
@

>>>>

<<trndmap, fig.cap = 'Percent change per year in nitrogen analytes for (a) annual and (b) seasonal aggregations.  Changes are based on seasonal Kendall tests of flow-normalized results within each time period. Station names are shown in the top left panel. Three values less than 5\\% change were truncated for visual clarity of the bar plots, see \\cref{tab:trnds} for actual values. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.', fig.height = 8, fig.width = 7, cache = T, out.width = '\\textwidth'>>=

res = c('din', 'nh', 'no23')
mobrks <- list(c(3, 4, 5), c(6, 7, 8), c(9, 10, 11), c(12, 1, 2))
yrbrks <- c(-Inf, 1995, Inf)
molabs <- c('Spring', 'Summer', 'Fall', 'Winter')
yrlabs <- c('1976-1995', '1996-2014')
buffx = 0.00035
buffy = 0.0022
cols =  c('#D53E4F', '#4DAF4A')
sz_rng = c(2, 11)
shps = c(24, 25)
ndivs = 6
strp_fl = 'grey90'
bar = FALSE
stat_labs = TRUE
leg = TRUE
leg_lab = 'Percent change'

# load required data
data(delt_dat)
data(delt_map)
data(trnds_sk)

##
# get a bounding box for the stations
statmeta <- select(delt_dat, Site_Code, Latitude, Longitude) %>% 
  unique

# bounding box
lims<- select(statmeta, -Site_Code) %>% 
  apply(., 2, function(x) range(x, na.rm = TRUE))
lims <- c(
  (1 + buffx) * lims[1, 2], 
  (1 - buffy) * lims[1, 1], 
  (1 - buffx) * lims[2, 2], 
  (1 + buffy) * lims[2, 1]
  )    
names(lims) <- c('left', 'bottom', 'right', 'top')

# response labels 
reslabs <- c('din an', 'nh an', 'no23 an', 'din mo', 'nh mo', 'no23 mo')
reslevs <- c(
  'DIN', 
  expression(paste(NH [4] ^ '+')),
  expression(paste(NO [2] ^ '-', ' / ', NO [3] ^ '2-')),
  'DIN', 
  expression(paste(NH [4] ^ '+')),
  expression(paste(NO [2] ^ '-', ' / ', NO [3] ^ '2-'))
)

# convert delta map to ggplot compatible object
delt_map <- fortify(delt_map)

# base delta map
alph <- 1
pbase <- ggplot(delt_map, aes(x = long, y = lat)) + 
  geom_polygon(aes(group = group, fill = hole), colour = alpha("lightblue", alph)) +
  theme_bw() + 
  theme(
    plot.background = element_rect(fill='transparent', 
      colour = NA),
    panel.background = element_rect(fill='transparent', 
      colour = NA),
    legend.background = element_rect(fill='transparent', 
      colour = NA),
    legend.key = element_rect(fill = 'transparent', 
      colour = NA),
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"), 
    panel.grid.minor = element_blank(), 
    axis.title.x = element_blank(), 
    axis.title.y = element_blank(), 
    legend.position = 'none', 
    plot.margin = grid::unit(c(3, 3, 3, 3), 'pt'), 
    axis.text.x = element_text(size = 8), 
    axis.text.y = element_text(size = 8), 
    strip.background=element_rect(fill = 'white', colour = 'white'),
    panel.border = element_rect(colour = "grey")
  ) +
  coord_fixed(ratio = 1, xlim = lims[c(1, 3)], ylim = lims[c(2, 4)]) +
  scale_fill_manual(values=c(alpha("lightblue", alph), "white"), guide="none")

yrng <- c(-5, 5)

# color ramp, red for pos, green for neg
colpos <- RColorBrewer::brewer.pal(9, 'Reds')[3:9]
colneg <- RColorBrewer::brewer.pal(9, 'Greens')[3:9]
palneg <-  colorRampPalette(colneg)(1000)
palpos <-   colorRampPalette(colpos)(1000)

# plot(1:2000, col = c(palneg, palpos), pch = 16, cex = 4)
# points(1000, 1000)

# unnest data frame, assign colors by rank and sign
toplo <- select(trnds_sk, Site_Code, resvar, resvar, cat, ztest, perchg) %>% 
  left_join(., statmeta, by = 'Site_Code') %>% 
  data.frame %>% 
  mutate(
    perchg = pmax(yrng[1], perchg),
    perchg = pmin(yrng[2], perchg),
    negs = sign(perchg),  
    rnks = scales::rescale(abs(perchg), c(1, 1000)), 
    rnks = round(rnks), 
    cols = ifelse(negs == 1, palpos[rnks], palneg[rnks]),
    ind = 1:nrow(.), 
    cattype = ifelse(cat %in% molabs, 'mo', 'an')
  )
# plot(perchg ~ ind, dat = toplo, col = cols)

unis <- with(toplo, paste(resvar, cattype)) %>% 
  unique %>% 
  sort

sites <- unique(toplo$Site_Code)

plo_ls <- vector('list', length = length(unis))
names(plo_ls) <- unis

for(uni in unis){

  # get variable, agg period
  splits <- strsplit(uni, ' ')[[1]]
  resvart <- splits[1]
  cattypet <- splits[2]
  toplo_tmp <- filter(toplo, resvar %in% resvart & cattype %in% cattypet)

  # create individual bar plots as ggplot list
  bars <- setNames(lapply(sites, function(i){

    subplo <- filter(toplo_tmp, Site_Code %in% i)
    ggplot(subplo, aes(x=cat, y = perchg, fill = factor(cat))) +
      geom_bar(stat = 'identity', colour = 'black', size = 0.2) + 
      scale_fill_manual(values = as.character(subplo$cols)) + 
      theme_tree() + 
      xlab(NULL) + 
      ylab(NULL) + 
      theme_transparent() +
      theme(plot.margin=unit(c(0,0,0,0),"mm")) + 
      scale_y_continuous(limits = yrng)

  }), sites)

  ##
  # add the barplots to the base map

  # add site labels if plot is first
  pall <- pbase
  set.seed(5)
  if(which(uni == unis) == 1)
    pall <- pbase + 
      geom_label_repel(
        data = statmeta, 
        aes(x = Longitude, y = Latitude, label = Site_Code), 
        label.r = unit(0, "lines"),
        box.padding = unit(2, "lines"),
        label.padding = unit(0.15, "lines"), 
        point.padding = unit(0.1, "lines"),
        force = 2, size = 2,
        fill = strp_fl
      )

  # keep bar widths the same for cattypes
  subwid <- 0.1
  if(grepl('an$', uni))
    subwid <- subwid/2

  for(i in 1:length(bars)){

      nms <- names(bars)[i]
      dat <- statmeta[statmeta$Site_Code %in% nms,]
      pall <- subview(pall, bars[[i]], x = dat$Longitude, y = dat$Latitude, subwid, 0.14)

  }

  tval <- reslevs[which(uni == reslabs)]
  pall <- pall + 
    geom_text(data = data.frame(x = lims[1], y = lims[4], tval = as.character(tval)), 
      aes(x = x, y = y, label = tval), parse = TRUE, hjust = 0, vjust = 1)
  plo_ls[[uni]] <- pall

}

# setup legend
leg <- ggplot(data.frame(x = 1:10, y = 1:10, z = seq(-5, 5, length = 10)), aes(x, y, colour = z)) + 
  geom_point() + 
  scale_colour_gradientn(colours = c(rev(colneg), colpos)) + 
  guides(colour = guide_colourbar(title = expression('% chg yr' ^-1), barheight = 5, barwidth = 0.5))
leg <- g_legend(leg)

grid.arrange(
  arrangeGrob(plo_ls[[1]], plo_ls[[3]], plo_ls[[5]], ncol = 1, top = c('Annual (1976-1995, 1996-2014)')), 
  arrangeGrob(plo_ls[[2]], plo_ls[[4]], plo_ls[[6]], ncol = 1, top = c('Seasonal (Sp, Su, Fa, Wi)')),
  leg,
  ncol = 3, 
  widths = c(1, 1, 0.4)
  )

### 2016-11-21 sftrends_manu.tex
working on subplot map
working on barplot map
new trends map with bar plots is done

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$, Phil Trowbridge$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Fri Nov 18 13:04:31 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

{\centering \includegraphics[width=0.75\textwidth]{figs/trndmap-1} 
\includegraphics[width=0.75\textwidth]{figs/trndmap-2}

@@

\caption[Percent changes in nitrogen analytes for (a) annual and (b) seasonal (monthly) periods in the record]{Percent changes in nitrogen analytes for (a) annual and (b) seasonal (monthly) periods in the record.  Changes are based on the difference between the ending and starting estimates for the flow-normalized estimates within each period.  Points are colored for direction (red increasing, green decreasing) and sized for relative magnitude. Station names are shown in the top left panel. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.}\label{fig:trndmap}
\end{figure}

>>>>

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$, Phil Trowbridge$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Sun Nov 20 17:44:15 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

{\centering \includegraphics[width=\textwidth]{figs/trndmap-1}

@@

\caption{Percent change per year in nitrogen analytes for (a) annual and (b) seasonal aggregations.  Changes are based on seasonal Kendall tests of flow-normalized results within each time period. Station names are shown in the top left panel. Three values less than 5\% change were truncated for visual clarity of the bar plots, see \cref{tab:trnds} for actual values. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.}\label{fig:trndmap}
\end{figure}
